Results

### 2022-05-15 article/abs.Rmd
draft abs esp

>>>>

# Esp

Más de la mitad de bosques son tropicales y secundarios, indicando que la regeneración de bosques tropicales es integral al futuro de bosques, globalmente.
Biomasa arboreal y *riqueza* taxonomica pueden recuperarse en algunas décadas, pero abundancias relativas pueden durar más de un siglo.
Dado que la mayoría de bosques están entre un km de su orilla, *efectos de orilla* probablemente afectan la regeneración de composición de comnuidad, mediado por tolerancia a sombra y "traits"/characterísticas de *disperso*, y también historia de manejamiento.
Este estudio *investiga* cómo regeneración de bosque secundario tropical húmedo es afectado por distancia a *orilla de habitat*, y *"hypothesizes"* que plantaciones de madera viejos facilitan regeneración porque favoran taxa tolerante a sombra o de *"succession"* tarde, y que orillas de bosque cercanos/*"adjacent"* *median* composición de comunidad.
Una plantación de madera neotropical *al lado de* bosque primario fue *"censused"* por árboles siguiendo un transecto de 300 m midiendo un gradiente de distancia de orilla, y análisis combino taxa con characterísticas/*"traits"* usando literatura relevante.

Mientras distancia de bosque primario aumentó, densidad de madera y *"stems"* *"tended to"* aumentó *signifantemente*, con 10% variación explicado, mientras tanto biomasa y luz de dócel *"tended to"* quedarse igual.
*Riquesa* de árboles también *"tended to"* aumentar *significantemente*, pero diversidad bajó profundamente y no-linearmente, y composición taxonomica varió notablemente.
Finalmente, taxa de árboles asociado *etapas* de *"succession"* ambos temprano y tarde  bajó *significativamente*, pero biomasa por modo de *disperso* no *"tended to"* cambiar.
*Entre todo* este estudio apoya que composición es menos resiliente y más sujeto a *efectos de orilla* que biomasa y *riqueza*, sugeriendo que bosques globales probablemente serán *asemblajes* nuevos distintos en el futuro, con *"trade-offs"* entre producción de madera y biodiversidad *ocuriendo* basado en actividad de manejamiento local y regional.

### 2022-05-15 article/methods.Rmd
draft abs esp

For all response variables, data were analyzed using plot medians and subject to linear regression with the distance to forest edge variable after checks for residual normality and equal variance assumptions using `r version$version.string`.

 <!-- including biomass converted to a per hectare basis,  -->

<!-- Correlated variables were grouped into multi-way ANOVAs---namely biomass, wood density, and diameter in one group; and height, stem density, and wood density in a different group.  -->

Non-linear regressions were run using the function *poly()* in the *stats* `r packageVersion("stats")` R package *(R Core Team 2022)*, 
which was only recorded for stem density and richness variables. 
All trees censused were included in each analysis including unidentified taxa that represented a single group in taxa-specific analyses.
Community analysis and ordination was done as a PERMANOVA with the *vegan* `r packageVersion("vegan")` R package *[@dixon03]*.
Data and code are available at [nmedina17.github.com/osa](nmedina17.github.com/osa).

>>>>

For all response variables, data were analyzed using plot medians and subject to linear regression with distance to forest edge as the only independent variable, after checks for residual normality and equal variance assumptions, using `r version$version.string`.
 <!-- including biomass converted to a per hectare basis,  -->

<!-- Correlated variables were grouped into multi-way ANOVAs---namely biomass, wood density, and diameter in one group; and height, stem density, and wood density in a different group.  -->
Non-linear regressions were run using the function *poly()* in the *stats* `r packageVersion("stats")` R package *(R Core Team 2022)*, 
which was only recorded for stem density and richness variables. 
All trees censused were included in each analysis including unidentified taxa that represented a single group in taxa-specific analyses.
Community analysis and ordination was done as a PERMANOVA with the *adonis()* function from the *vegan* `r packageVersion("vegan")` R package *[@dixon03]*.
Data and code are available at [nmedina17.github.com/osa](nmedina17.github.com/osa).

### 2022-05-15 article/results.Rmd
draft abs esp

---
title: "Results"
output: html_notebook
---

```{r setup, echo = F, message = F}
knitr::opts_chunk$set(
  echo = F,
  message = F,
  warning = F
)
library(here)
```

## Biomass

Distance to primary forest edge was associated with ### more ___, although this did not correlate with ___.

@@

```{r fig1}
source(
  here(
    "figs/fig1.R"
  )
)
fig1
```

## Composition

@@

```{r fig2}
source(
  here(
    "figs/fig2.R"
  )
)
fig1
```

>>>>

---
title: "Results"
output: 
  # html_notebook
  bookdown::html_document2:
    number_sections: FALSE
---

```{r setup, echo = F, message = F, warning=F, results=F}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
here::i_am("article/results.Rmd"); source(here::here("figs/fig2.R"))
```

# Biomass

@@

```{r fig1data}
wd <- filter(plotResultsTbl1, variable == "spg") #alias
stems <- filter(plotResultsTbl1, variable == "stems") #alias
stems12 <- filter(plotResultsTbl12, variable == "stems") %>% distinct() #alias
```

Aboveground biomass did not tend to change with distance to primary forest edge 
(Fig \@ref(fig:biomass)a), despite significant trends with underlying related variables 
(Fig \@ref(fig:biomass)).

Median plot wood density tended to increase marginally significantly 
(`r glue("p=",round(as.numeric(wd$pval),3))`) 
with edge distance by 
`r round(wd$statPrint[[1]]$estimate[[2]],4)` ± 
`r round(wd$statPrint[[1]]$std.error[[2]],5)` per m 
explaining 
`r "R2x100"`% of variance 
among median distance strata values
(Fig \@ref(fig:biomass)b).

Plot stem density also tended to increase significantly 
(`r glue("p=",round(as.numeric(stems12$pval[[2]][2]),2))`) 
with edge distance but more strongly and non-linearly, by 
~`r round(stems12$statPrint[[1]]$estimate[[2]],2)` ± 
`r round(stems12$statPrint[[1]]$std.error[[2]],2)` per m 
(Fig \@ref(fig:biomass)c).

At this stage in forest regeneration, canopy light availability did not tend to change with distance from forest edge 
(Fig \@ref(fig:biomass)d).

```{r biomass, fig.cap=" Plot stand properties, namely (a) biomass, (b) wood density, (c) stem density, and (d) canopy light availability."}
knitr::include_graphics(here::here("../figs/fig1.png"))
```

# Composition

@@

```{r fig2data}
entropy <- filter(plotResultsTbl1, variable == "entropy") #pReplaced
entropy12 <- filter(plotResultsTbl12, variable == "entropy") %>% distinct() #slopeR2
richness <- filter(plotResultsTbl1, variable == "richness")
ficus <- filter(taxaResultsTbl1, gen == "Ficus")
```

Overall diversity changed significantly with distance to primary forest edge 
(Fig \@ref(fig:taxa)).

Alpha diversity significantly 
(`r glue("p=",round(as.numeric(entropy12$pval[[2]][2]),3))`) 
decreased non-linearly by 
~`r round(abs(entropy12$statPrint[[1]]$estimate[[2]]),2)` ± 
`r round(abs(entropy12$statPrint[[1]]$std.error[[2]]),2)` per m 
as distance to primary forest edge increased 
which explained 
`r "R2x100"`% of variance 
among distance strata medians 
(Fig \@ref(fig:taxa)a) -- 
while in contrast taxonomic richness increased slightly 
(`r glue("p=",round(as.numeric(richness$pval),3))`) 
and linearly 
by 
`r round(richness$statPrint[[1]]$estimate[[2]],2)` ± 
`r round(richness$statPrint[[1]]$std.error[[2]],3)` per m 
and had `r "R2x100"`% variance explained 
among distance strata 
(Fig \@ref(fig:taxa)b).

Community composition and beta diversity also changed significantly 
(`r glue("p=",round(ordStat$aov.tab[[6]][1],2))`) with 
`r round(ordStat$aov.tab$R2[1],2) *100`% 
variance among distance strata explained by distance to primary forest edge, 
and the first two principal components explaining 
`r ordTbl$PC1_prop[1] *100`% and 
`r ordTbl$PC2_prop[1] *100`% totaling 
`r (ordTbl$PC1_prop[1] + ordTbl$PC2_prop[1]) *100`% 
of variance among calculated plot distances 
(Fig \@ref(fig:taxa)d).

Key abundant taxa `r measure2e` and `r measure2f` showed different responses -- `r measure2e` nearly tended to decrease with distance to primary forest edge 
(Fig \@ref(fig:taxa)e),
while `r measure2f` decreased marginally significantly 
(`r glue("p=",round(as.numeric(ficus$pval),3))`) 
and linearly by 
`r round(ficus$statPrint[[1]]$estimate[[2]],2)` ± 
`r round(ficus$statPrint[[1]]$std.error[[2]],2)` per m 
with `r "R2x100"`% variance among distance strata explained 
(Fig \@ref(fig:taxa)f).

```{r taxa, fig.cap=" Taxonomic composition measures including (a) diversity, (b) richness, (c) overall biomass, (d) beta diversity, and (e,f) biomass of key taxa."}
knitr::include_graphics(here::here("../figs/fig2.png"))
```

# Functions

```{r fig3data}
source(here::here("figs/fig3.R")); succession <- taxaResultsTbl12 #pReplaced
```

Overall functional regeneration highlighted successional stage associations over primary dispersal mode along distance to primary forest 
(Fig \@ref(fig:traits)).

Taxa associated with both early and late successional stages decreased significantly 
(`r glue("p=",round(as.numeric(succession$pval[[3]]),3))`) 
and curvi-linearly 
by 
~`r round(succession$statPrint[[1]]$estimate[[2]],2)` ± 
`r round(succession$statPrint[[1]]$std.error[[2]],2)` per m 
with distance to edge explaining 
`r "R2x100"`% variance 
among distance strata medians 
(Fig \@ref(fig:traits)a).

Dispersal modes did not show consistent trends in biomass with increasing distance to edge 
(Fig \@ref(fig:traits)b).

```{r traits, fig.cap=" Functional regeneration based on taxa associations with (a) successional stages and (b) primary dispersal mode."}
knitr::include_graphics(here::here("../figs/fig3.png"))
```
